---
title: "Getting started with wbcmd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with wbcmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

If you work with neuroimaging data in R, you've probably typed out a `system()` call to `wb_command` at some point.
Maybe you wrapped it in a helper function.
Maybe that helper grew.
wbcmd is the version of that helper that validates your inputs, finds your binary, and gets out of the way.

## Finding the binary

Install [Connectome Workbench](https://www.humanconnectome.org/software/connectome-workbench), then load the package:

```{r}
#| label: check
library(wbcmd)
wb_sitrep()
```

`wb_sitrep()` checks whether the binary is reachable and prints what it finds.
On macOS and Linux, wbcmd searches common installation paths and the system `PATH` automatically.

If the binary lives somewhere unusual, tell wbcmd where:

```{r}
#| label: set-path
set_wb_path("/opt/workbench/bin_linux64/wb_command")
```

The full lookup order is:

1. R option `wbcmd.path`
2. Environment variable `WB_PATH`
3. Platform-specific defaults (e.g. `/Applications/wb_view.app/...` on macOS)
4. System `PATH`

Set `WB_PATH` in your `.Renviron` and you never think about it again.

## The direct route

`wb_cmd()` is the universal entry point.
It takes the subcommand name and arguments exactly as you would type them on the command line:

```{r}
#| label: wb-cmd
wb_cmd("-version")
wb_cmd("-file-information", "data.dscalar.nii")
```

Every wrapper function in wbcmd calls `wb_cmd()` internally.
If a wrapper doesn't exist for the command you need, `wb_cmd()` always works.

## Wrappers

The wrappers handle argument construction, file validation, and output path management.
They error before the command runs if an input file is missing, and they create output directories on the fly.

### Parcellate CIFTI data

```{r}
#| label: parcellate
cifti_parcellate(
  "data.dtseries.nii",
  "atlas.dlabel.nii",
  direction = "COLUMN",
  cifti_out = "parcellated.ptseries.nii"
)
```

### Compute a correlation matrix

```{r}
#| label: correlation
cifti_correlation(
  "data.dtseries.nii",
  "corr.dconn.nii",
  fisher_z = TRUE
)
```

### Reduce across maps

```{r}
#| label: reduce
cifti_reduce(
  "data.dtseries.nii",
  direction = "ROW",
  operation = "MEAN",
  cifti_out = "mean.dscalar.nii"
)
```

### Surface curvature

```{r}
#| label: curvature
surface_curvature(
  "lh.midthickness.surf.gii",
  mean_out = "lh.curvature.func.gii"
)
```

### Map a volume to a surface

```{r}
#| label: vol-to-surf
volume_to_surface_mapping(
  "bold.nii.gz",
  "lh.midthickness.surf.gii",
  metric_out = "lh.bold.func.gii",
  method = "ribbon-constrained",
  inner_surface = "lh.white.surf.gii",
  outer_surface = "lh.pial.surf.gii"
)
```

### Math expressions

`cifti_math()`, `metric_math()`, and `volume_math()` all take a symbolic expression and a named list of file paths:

```{r}
#| label: math
cifti_math(
  "(x + y) / 2",
  cifti_out = "average.dscalar.nii",
  var = list(
    x = "subj01.dscalar.nii",
    y = "subj02.dscalar.nii"
  )
)
```

### Getting help

Every wrapper function includes a "Connectome Workbench Help" section in its documentation, pulled directly from `wb_command` at build time.
You can also call `wb_help()` interactively:

```{r}
#| label: help
wb_help("-cifti-parcellate")
```

## Using wbcmd with ciftiTools

[ciftiTools](https://github.com/mandymejia/ciftiTools) reads and writes CIFTI files as R objects.
wbcmd runs Workbench operations on those files.
They complement each other — ciftiTools gives you the data in R, wbcmd gives you the processing pipeline.

Both packages need the same binary.
Point them to the same path and they share the setup:

```{r}
#| label: ciftitool-setup
wb_path <- "/opt/workbench/bin_linux64/wb_command"

wbcmd::set_wb_path(wb_path)
ciftiTools::ciftiTools.setOption("wb_path", wb_path)
```

A typical workflow reads data with ciftiTools, processes files with wbcmd, then reads the results back:

```{r}
#| label: ciftitools-workflow
library(ciftiTools)
library(wbcmd)

xii <- read_cifti("data.dtseries.nii", brainstructures = "all")

cifti_parcellate(
  "data.dtseries.nii",
  "atlas.dlabel.nii",
  direction = "COLUMN",
  cifti_out = "parcellated.ptseries.nii"
)

parcellated <- read_cifti("parcellated.ptseries.nii")
```

ciftiTools also has its own `run_wb_cmd()` function that takes a pre-formatted command string.
If you already use that, wbcmd's `wb_cmd()` does the same job with the addition of path detection, argument quoting, and verbose logging.

## Controlling verbosity

By default, wbcmd prints the full command before running it.
Turn that off for cleaner output:

```{r}
#| label: verbose
options(wbcmd.verbose = FALSE)
```

Or set `WB_VERBOSE=FALSE` in your `.Renviron`.
